Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations

### 2022-11-04 paper/paper/SPIE/article.tex
[WIP]

In order not to inundate readers with dataset details, and because this is an
exploratory investigation of RMT which is not committed to any specific
theory\footnote{In fact, datasets and, later, the classification tasks, were
chosen without investigation into any findings or publications associated with
the data, both in order to reduce bias in our decisions, and because the
original authors analytic / statistical approaches and intentions have very
limited relevance to our whole-brain, voxelwise, multiverse predictive
approach.}, we refer the reader to the original publications and/or data
releases when greater detail is needed, and highlight only the most basic
aspects of each dataset here.

Dataset scan parameters and acquisition details are summarized in Table
\ref{table:1}, and sample and subgroup sizes are summarized in Table
\ref{table:2}.

@@

@@

\begin{table}[h!]
\caption{
    ID = Identifier for paper. TR = Time of Repetition (seconds).
    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
    indicate P horizontal / axial slices each with dimensions M x N.
}
\label{table:1}
\small
\centering
\begin{tabular}{ c c c c c c c }
\hline
\textbf{ID}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
\hline
AGING     & 74 x 74 x 32   & 3.0 x 3.0 x 4.0 & 2.0  & 300 & 62  \\
BILING    & 100 x 100 x 72 & 1.8 x 1.8 x 1.8 & 0.88 & 823 & 90  \\
DEPRESS   & 112 x 112 x 25 & 2.0 x 2.0 x 5.0 & 2.5  & 100 & 72  \\
LEARN     & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0  & 195 & 432 \\
OSTEO     & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5  & 300 & 74  \\
PARK      & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4  & 149 & 552 \\
PSYCH     & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0  & 300 & 90  \\
\hline
\end{tabular}

@@

\end{table}

\begin{table}[h!]
\caption{
    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
}
\label{table:2}
\small
\centering
\begin{tabular}{ l l c c c }
\hline
\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
\hline
AGING   &  older             & 28      & resting-state         & 1 \\
        &  younger           & 34      & resting-state         & 1 \\
\hline
BILING  &  bilingual          & 59      & resting-state        & 1 \\
        &  monolingual        & 33      & resting-state        & 1 \\
\hline
DEPRESS &  depression         & 51      & resting-state        & 1 \\
        &  control            & 21      & resting-state        & 1 \\
\hline
LEARN   &  task              & 24      & learn image sketches  & 16 \\
        &  rest              & 24      & resting-state         & 2 \\
\hline
OSTEO   &  duloxetine        & 19      & taking duloxetine     & 1 \\
        &  pain              & 37      & placebo               & 1 \\
        &  nopain            & 20      & control               & 1 \\
\hline
PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
        &  control           & 21      & ANT                   & 12 \\
\hline
PSYCH   &  vigilant           & 11     & resting-state        & 3 \\
        &  nonigilant         & 11     & resting-state        & 3 \\
        &  trait-attentive     & 11    & resting-state        & 3 \\
        &  trait-non-attentive & 11    & resting-state        & 3 \\
        &  task-attentive     & 11     & resting-state        & 3 \\
        &  task-nonattentive  & 11     & resting-state        & 3 \\
\hline
\end{tabular}
\end{table}

>>>>

In order not to inundate readers with dataset details, and because this is an
exploratory investigation of RMT which is not committed to any specific
theory\footnote{In fact, datasets and, later, the classification tasks, were
chosen without investigation into any findings or publications associated with
the data, both in order to reduce bias in our decisions, and because the
original authors analytic / statistical approaches and intentions have very
limited relevance to our whole-brain, voxelwise, multiverse predictive
approach.}, we refer the reader to the original publications and/or data
releases when greater detail is needed, and highlight only the most basic
aspects of each dataset here.  Dataset scan parameters and acquisition details
are summarized in Table \ref{table:1}, and sample and subgroup sizes are
summarized in Table \ref{table:2}.

\begin{table}[h!]
\caption{
    ID = Identifier for paper. TR = Time of Repetition (seconds).
    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
    indicate P horizontal / axial slices each with dimensions M x N.
}
\label{table:1}
\small
\centering
\begin{tabular}{ l c c c c c c }
\hline
\textbf{Name}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
\hline
Aging     & 74 × 74 × 32   & 3.0 × 3.0 × 4.0 & 2.0  & 300 & 62  \\
Bilingual    & 100 × 100 × 72 & 1.8 × 1.8 × 1.8 & 0.88 & 823 & 90  \\
Depress   & 112 × 112 × 25 & 2.0 × 2.0 × 5.0 & 2.5  & 100 & 72  \\
Learn     & 64 × 64 × 36   & 3.0 × 3.0 × 3.0 & 2.0  & 195 & 432 \\
Osteo     & 64 × 64 × 36   & 3.4 × 3.4 × 3.0 & 2.5  & 300 & 74  \\
Park      & 80 × 80 × 43   & 3.0 × 3.0 × 3.0 & 2.4  & 149 & 552 \\
Attention & 128 × 128 × 70 & 1.5 × 1.5 × 1.5 & 3.0  & 300 & 90  \\
\hline
\end{tabular}
\end{table}

@@

The \textit{Depression} dataset\cite{bezmaternykhddRestingStateClosed2021,
bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans frmo
non-depressed controls and mildy- or moderately-depressed subjects. The our analysis,
the binary classification task is \code{depress v control}.

% The intergroup study involved 21 healthy volunteers received a compensation
% for a participation and 51 participants featuring mild depression (F32.0),
% moderate depression (F32.1), or dysthymia (F34.1, single patient) (see Table
% 1). The fMRI study was carried out in the International Tomography Center,
% Novosibirsk, using a 3 T Ingenia scanner (Philips). Functional T2∗-weighted
% Ssh echo planar imaging scans were acquired using the following parameters:
% voxel size 2×2×5mm, repetition time/echo time = 2500/35 ms, and fat
% suppression mode. The reference anatomical image was obtained by the T1W 3D
% turbo field echo method with a voxel size of 1×1×1mm. The instruction for
% participants was to lie still with eyes closed for 6 minutes.

@@

@@

\begin{table}[h!]
\caption{
    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
}
\label{table:2}
\small
\centering
\begin{tabular}{ l l c c c }
\hline
\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
\hline
AGING   &  older             & 28      & resting-state         & 1 \\
        &  younger           & 34      & resting-state         & 1 \\
        &                    &         &                       &   \\
BILING  &  bilingual          & 59      & resting-state        & 1 \\
        &  monolingual        & 33      & resting-state        & 1 \\
        &                    &         &                       &   \\
DEPRESS &  depression         & 51      & resting-state        & 1 \\
        &  control            & 21      & resting-state        & 1 \\
        &                    &         &                       &   \\
LEARN   &  task              & 24      & learn image sketches  & 16 \\
        &  rest              & 24      & resting-state         & 2 \\
        &                    &         &                       &   \\
OSTEO   &  duloxetine        & 19      & resting-state         & 1 \\
        &  pain              & 37      & resting-state         & 1 \\
        &  nopain            & 20      & resting-state         & 1 \\
        &                    &         &                       &   \\
PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
        &  control           & 21      & ANT                   & 12 \\
        &                    &         &                       &   \\
PSYCH   &  vigilant           & 11     & resting-state        & 3 \\
        &  nonigilant         & 11     & resting-state        & 3 \\
        &  trait-attentive     & 11    & resting-state        & 3 \\
        &  trait-non-attentive & 11    & resting-state        & 3 \\
        &  task-attentive     & 11     & resting-state        & 3 \\
        &  task-nonattentive  & 11     & resting-state        & 3 \\
\hline
\end{tabular}
\end{table}
